---
title: "Case Study 8"
author: "Claire Hays"
date: "10/25/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Packages}

library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(lubridate)
library(ggplot2)
library(grid)
library(corrplot)
library(COR)
library(readr) 
library(haven)
library(readxl)
library(foreign)
library(stringr)
library(stringi)

```

```{r read in data}
sales_dat <- tempfile()
download("https://github.com/WJC-Data-Science/DTS350/raw/master/sales.csv", sales_dat, mode = "wb")
dat <- read_csv(sales_dat)
head(dat)
str(dat)
summary(dat)
```

```{r change timezone}
sales <- with_tz(dat, tzone = "US/Mountain")
```

```{r time aggregations}
sales_time <- sales %>%
  mutate(Month = month(Time, label = TRUE, abbr = FALSE)) %>%
  mutate(Week = week(Time)) %>%
  mutate(Day = wday(Time, label = TRUE, abbr = FALSE)) %>%
  mutate(Hour = hour(Time)) 

head(sales_time)
tail(sales_time)
summary(sales_time)
```

```{r reccomendations}
HOURS <- sales_time %>%
  group_by(Name, Hour) %>%
  select(Name, Amount, Hour) %>%
  ggplot(aes(x=Hour, y=Amount, Color=Name)) +
  geom_col() +
  facet_wrap(~Name) +
  labs(title = "Amount Sold per Hour for Each Store",
       x = "Hour of Day",
       y = "Amount Sold") +
  theme_bw()


DAYS <- sales_time %>%
  group_by(Name, Day) %>%
  select(Name, Amount, Day) %>%
  ggplot(aes(x=Day, y=Amount, Color=Name)) +
  geom_col() +
  facet_wrap(~Name) +
  labs(title = "Amount Sold per Day of Week for Each Store",
       x = "Day of Week",
       y = "Amount Sold") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


MONTHS <- sales_time %>%
  group_by(Name, Month) %>%
  select(Name, Amount, Month) %>%
  ggplot(aes(x=Month, y=Amount, Color=Name)) +
  geom_col() +
  facet_wrap(~Name) +
  labs(title = "Amount Sold per Month for Each Store",
       x = "Month",
       y = "Amount Sold") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
```
I would recommend hours of operation be between 8am and 9pm, with customer traffic being the busiest between 10am and 4pm for just about every restaruant.  I also reccomend operating during the weekdays.  Weekend sales are low amongst all stores.  We are only comparing a few months of the year, but it does appear the June is the busiest month for almost all the sotres (except LeBelle). 

```{r final comparison}
COMPARE <- sales_time %>%
  group_by(Name) %>%
  ggplot(aes(x=Name, y=Amount, Color=Name)) +
  geom_col() +
  labs(title = "Sales Across All Stores",
       x = " ",
       y = "Sales") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

I reccomend waiting to open Frozone until June.