---
title: "Task 12"
author: "Claire Hays"
date: "10/5/2022"
output: 
  html_document:
    theme: cosmo
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(lubridate)
library(ggplot2)
library(grid)
library(corrplot)
library(COR)
library(readr) 
library(haven)
library(readxl)

```

```{r files}

dat_1 <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.RDS", dat_1)
dat1 <- read_rds(dat_1)
head(dat1)

dat_2 <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.csv", dat_2)
dat2 <- read_csv(dat_2)
head(dat2)

dat_3 <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.dta", dat_3)
dat3 <- read_dta(dat_3)
head(dat3)

dat_4 <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.sav", dat_4)
dat4 <- read_sav(dat_4)
head(dat4)

dat_5 <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.xlsx", dat_5)
dat5 <- read_xlsx(dat_5)
head(dat5)

all_equal(dat1, dat2, dat3, convert = TRUE)
all_equal(dat1, dat4, dat5, convert = TRUE)
```

```{r graphic}

boxplot <- ggplot(data=dat1, aes(x=variable, y=value, fill=variable)) +
           geom_boxplot()
boxplot



jitter <- ggplot(data=dat1, aes(x=variable, y=value, fill=variable)) +
           geom_jitter()
jitter



average_return <- ggplot(data=dat1, aes(x=contest_period, y=value, group=variable, color=variable)) +
                  geom_line() +
                  facet_wrap(~variable)
average_return
```
boxplot - PROS has the best potential and also the most risk.  DJIA doesn't have as high of potential as PROS but it is "safer" because there is not as much margin.  DARTS is reasonably "safe" but the potential is not as significant as PROS.  

jitter - mostly similar to what the boxplot showed us.  this jitter plot does allow us to see that there are few outliers for DARTS, allowing there to be potentional a gain almost as high as we see using PROS.  

average_return - DJIA would be the easiest to predict.  DART and PROS have much more variation.

```{r tidy}
dat1_tidy <- dat1 %>%
             filter(variable == 'DJIA') %>%
      mutate(contest_period = str_replace_all(contest_period, '19', '_19'),
             contest_period = str_replace(contest_period, 'Febuary', 'February'),
             contest_period = str_replace(contest_period, 'Dec._1993', 'December_1993')) %>%
      separate(contest_period, into = c('Month', 'Year_End'), sep = '-') %>%
      separate(Year_End, into = c('Month', 'Year'), sep = '_') %>%
      arrange(factor(Month, levels = month.name))

head(dat1_tidy, n=10)
saveRDS(dat1_tidy, file='dat1_tidy.rds')
```

```{r 6 month returns by year}
returns <- ggplot(data=dat1_tidy, aes(x=Year, y=value, group=Month, color=variable)) +
                  geom_line() +
                  geom_point() +
                  facet_wrap(~Month)
returns

```
returns - lots of variation among the different months.  October seems to have the least.  March and May also seem to have less than other months.  September and November have a lot of variation. April, February, May, and March all seemed to stay above 0.  

```{r Final Table}
final_table <- dat1%>%
  filter(variable == "DJIA") %>%
  mutate(contest_period = str_replace_all(contest_period, "19", "_19"),
         contest_period = str_replace(contest_period, "Febuary", "February"),
         contest_period = str_replace(contest_period, "Dec._1993", "December_1993")) %>%
  separate(contest_period, into = c("Month", "Year_end"), sep = "-") %>%
  separate(Year_end, into = c("Month", "Year"), sep = "_") %>%
  pivot_wider(names_from = "Year", values_from = "value") %>%
  arrange(factor(Month, levels = month.name)) %>%
  select(-variable)
head(final_table)
tail(final_table)

```



