---
title: "Task 7"
author: "Claire Hays"
date: "9/21/2022"
output: 
  html_document:
    theme: cosmo
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r Packages}
library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(ggplot2)
```


dowload data and create dataset
```{r Dataset}
getwd()
height_dat <- read_csv("heights.csv")
str(height_dat)

                         
                         
HeightEd <- height_dat %>%
  mutate(ed = fct_collapse(as.factor(ed),
                           primary = c("3","4","5","6"),
                           secondary = c("7","8","9","10","11"),
                           highschoolgrad = ("12"),
                           associate = c("13","14","15"),
                           bachelors = ("16"),
                           mastorsdoctorate = c("17","18")
                           ))
head(HeightEd)
tail(HeightEd)
```

relationship between education and race & education and sex
```{r Graphics}

ggplot(data = HeightEd) +
  geom_bar(mapping = aes(x = ed, fill = race), position = "dodge")

ggplot(data = HeightEd) +
  geom_bar(mapping = aes(x = ed, fill = sex), position = "dodge")

```


new graph
```{r New Graphic}
str(HeightEd)

ggplot(data = HeightEd, aes(x= race, y= earn, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y= "Earnings", x= "Race", title= "Earnings by Race and Sex") +
  annotate("text", x = "other", y = 130000, label = "Females much Higher", size = 3.5) +
  annotate("text", x = "black", y = 70000, label = "Similar", size = 3.5) +
  annotate("text", x = "hispanic", y = 70000, label = "Similar", size = 3.5) +
  annotate("text", x = "white", y = 220000, label = "Males much Higher", size = 3.5) +
  scale_fill_brewer(palette = "Dark2") +
  theme_bw()
  

```

We can see that among white females and males, overall, males make a much higher earning.  In the "other" category it found to be the opposite, with females making a notably higher earning.  Female and males among Black and hispanic races are found to make similar earnings. 

