---
title: "Case Study 12"
author: "Claire Hays"
date: "11/27/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(lubridate)
library(ggplot2)
library(grid)
library(corrplot)
library(readr) 
library(haven)
library(readxl)
library(foreign)
library(stringr)
library(stringi)
library(tidyquant)
library(timetk)
library(DT)
library(dygraphs)
library(sf)
library(USAboundaries)
library(USAboundariesData)
library(ggsflabel)
library(maps)
library(leaflet)
library(mapview)
library(gapminder)
library(gganimate)
library(gifski)
library(datasauRus)
```

```{r dowload data}
dat <- read_csv("avocado.csv")
head(dat)
summary(dat)
```

```{r separate region column into four sections}
region_column <- dat %>% distinct(region)
tail(region_column, 20)
head(region_column, 20)


total <- dat%>%
  filter(region == 'TotalUS')
regions <- dat %>%
  filter(region %in% c("GreatLakes", "MidSouth", "Northeast","NorthernNewEngland", "Plains", "SouthCentral", "Southeast","West", "WestTexNewMexico"))
states <- dat %>%
  filter(region == 'California')
cities <- dat %>%
  filter(! region %in% c("California","TotalUS", "GreatLakes", "MidSouth", "Northeast", "NorthernNewEngland", "Plains", "SouthCentral", "Southeast", "West", "WestTexNewMexico"))
```

```{r trends of totalUS}
prices_over_time <- ggplot(data=total, mapping = aes(x=Date, y =AveragePrice)) +
  geom_line() +
  labs(title = 'Avocado Price Trends Over The US',
       x = 'Date',
       y = 'Average Price') +
  theme_bw()
prices_over_time + transition_reveal(Date)
```

```{r avo prices over the regions}
names(regions)[names(regions) == "Total Volume"] <- "totalvolume"


price_over_regions <- ggplot(data=regions, mapping = aes(x=Date, y=AveragePrice, color = totalvolume)) +
  geom_line() +
  facet_wrap(~region) +
    labs(title = 'Avocado Price Trends Over US Regions',
       x = 'Date',
       y = 'Average Price') +
  theme_bw()
price_over_regions + transition_reveal(Date)

```

```{r avo prices over the cities} 
#show total volume with size = totalvolume
names(cities)[names(cities) == "Total Volume"] <- "totalvolume"

price_over_cities <- ggplot(data=cities, aes(x=totalvolume, y=AveragePrice, color=region)) +
  geom_line() +
  facet_wrap(~year) +
    labs(title = 'Avocado Price Trends Over Different Cities',
       x = 'Date',
       y = 'Average Price') +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
price_over_cities
```

```{r choose one: stocks}
start <- as.Date("2015-01-04")
end <- as.Date("2018-03-25")

CVGW <- tq_get(c("CVGW"), get = "stock.prices", from = start, to = end)
AVO <- tq_get(c("AVO"), get = "stock.prices", from = start, to = end)
#keep getting error message for AVO

CVGW_plot <- CVGW %>%  
          select(symbol, date, adjusted) %>%
          pivot_wider(names_from = symbol, values_from = adjusted) %>%
          tk_xts(date_var = date) 

dygraph(CVGW_plot, 
        main = "CVGW Adjusted Values") %>%
  dyAxis("y", label = "Adjusted")

```

For prices over time I used a geom_smooth to help my friend easily see the trends over the past several years. 
We see the total trends over the US is that prices have been fluctuating a lot but have increased over 2017-2018.
We see over the regions that avo prices have fluctuated a lot but have had a peak in most region in 2017-2018 like we saw in the overall price trends.  
When looking at the stock prices for the years 2015 - 2018 we can clearly see that they have been steadily increasing especially in the start of 2018.  
