---
title: "Task 10"
author: "Claire Hays"
date: "9/27/2022"
output: 
  html_document:
    theme: cosmo
    keep_md: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

```

```{r Packages}
library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(lubridate)
library(ggplot2)
library(grid)
library(COR)
library(corrplot)

```

```{r Explore}

devtools::install_github("drsimonj/ourworldindata")

getwd()
dat <- read_csv('financing_healthcare.csv')
str(dat)
head(dat)
tail(dat)

child <- dat %>%
  select(year, country, child_mort) %>%
  drop_na()
  

child_count <- child %>%
  select(country) %>%
  summarise(tot_country = n())

summary(child)
str(child)

child %>%
  count(country)
```

```{r Histogram}

ggplot(data = child, mapping = aes(x = year)) +
  geom_histogram(binwidth = 5)


```

I started by plotting a histogram to get a general idea of child mortality over the years.
There was almost no death from 1750 - 1800. This seemed bizarre to me, so I am going to look at this more.  From 1800 - 1930 it was relatively steady.  Around the 1950s there was a significant drip in mortality only for the numbers to increase until about 2000.  Entering into the 2000s, we see an extremely significant drop in child mortality.  


```{r 18th century}

century18 <- child %>%
  filter(year < 1800)

ggplot(data = century18, mapping = aes(x = year)) +
  geom_histogram(binwidth = 10)

ggplot() +
  geom_point(data = century18, aes(x = year, y = child_mort, color = country))



```
Now, we can see that these numbers were so low because the only country on record was Sweden.

```{r Just US}

child2 <- dat %>%
  select(year, country, child_mort, life_expectancy) %>%
  drop_na()
  
after1800 <- child2 %>%
  filter(year > 1800) %>%
  filter(country == 'United States')

summary(after1800)

ggplot() +
  geom_point(data = after1800, aes(x = year, y = life_expectancy, color = country)) +
  geom_point(data = after1800, aes(x = year, y = child_mort, color = country))


  
```
I decided to filter out the years before 1800 since those only had the data from Sweden. Then I decided to look just at the U.S. child_mortality versus life expectancy.  I plotted both these variable using a geom_point.  It seemed that these were negatively correlated, which would make since so I plotted a corrplot below.  

```{r Correlation of child_mort and life_expectancy in US}


justUS <- after1800 %>%
  select(life_expectancy, child_mort)

corrplot(cor(justUS), method = "ellipse",
                          type = "upper",
                          order = "hclust",
                          tl.col = "black",
                          tl.srt = 45) 


```

This corrplot confirms my expectation that in the U.S. as child mortality went down the life expectancy went up.  We can see this because there is a strongly negavtive correlation between the two variables.  