---
title: "Task 5"
author: "Claire Hays"
date: "9/12/2022"
output: 
  html_document:
    theme: cosmo
    keep_md: true
    code_following: 'hide'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(knitr)
library(downloader)

```


The data seemed to read in correctly as characters
```{r}

SoloData <- read_csv("solo-artist-followers.csv")
SoloData
str(SoloData)
```


The data seemed to read in correctly as characters, dates, and doubles.
```{r}

BillboardData <- read_csv("billboard-hits.csv")
str(BillboardData)
head(BillboardData)

```


remover all singers who did not hit top 100 6 times and find the corresponding bands
```{r}

sixtopsingers <- BillboardData %>%
  group_by(name) %>%
  filter(n() > 6, band != "", !is.na(band))
head(sixtopsingers)

sixtopbands <- BillboardData %>%
  group_by(band) %>%
  filter(band %in% sixtopsingers$band)
head(sixtopbands)

```


recronstruct the plot from task 5
oberservations:
In general, bands have alot of ups and downs.  Destiny's Child specifically shows a steady up, down, up, down, fluctuation over the years.  Tje Jonas brothers has a high peak around 2010 and then had a hard hit for several years (were the producing any songs then?), but then they had a few more succesfull songs in 2015 and 2017.    
```{r}

ggplot(data = sixtopsingers, aes(x = peak_date, y = peak_rank, color = name, group = name)) +
  geom_point() +
  geom_line() +
  geom_point(data = sixtopbands, color ="black") +
  geom_line(data = sixtopbands, color = "black", linetype = "dotted") +
  facet_wrap(~ band, scales = "free") +
  xlab("peak_date") + ylab("peak_rank") +
  theme_bw() 

```


Data comes from this source:
https://www.fbi.gov/services/cjis/nics

This source is realiable (FBI) information about Background Checks
All data read in as characters or column_doubles.
```{r}
FireArmDat <- read_csv("nics-firearm-background-checks.csv")
FireArmDat
str(FireArmDat)

```


Data comes from this source:
https://fivethirtyeight.com/features/should-travelers-avoid-flying-airlines-that-have-had-crashes-in-the-past/

This source was recommended (fivethirtyeight) and has a link to github with all the data used.
All data came in as characters or column_doubles.
```{r}
AirSafety <- read_csv("airline-safety.csv")
AirSafety
str(AirSafety)

```



Data comes from this source:
https://catalog.data.gov/dataset/?q=chronic+disease+indicators&page=2

This source was recommended and is a .gov source, so should be reliable.  It has useful data for discovreing how many students are physically fit.
All the data came in as a character or double.
```{r}

FitStudents <- read_csv("percentage-of-physically-fit-students-lghc-indicator-8.csv")
FitStudents
str(FitStudents)

```


