---
title: "Case Study 4"
author: "Claire Hays"
date: "9/21/2022"
output: 
  html_document:
    theme: cosmo
    keep_md: true
    code_folding: 'hide'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load packages
```{r Packages}
library(tidyverse)
library(knitr)
library(downloader)
library(dplyr)
library(ggplot2)
library(nycflights13)

```

QUESTION 2:
Which origin airport is best to minimize my chances of a late arrival when I am using Delta Airlines?
```{r filter data to answer question}
str(flights)
?flights

delta_airline<- filter(flights, carrier == "DL") %>%
  select(arr_delay, origin) %>%
  drop_na() %>%
  group_by(origin) %>%
  summarise(avg_delay = mean(arr_delay))



delta_airline

```
filter out delta airlines, then select the arrival delays and the origin to answer the question.
group by the origin to determine which origin of airport has the least likelyhood to have a delay.
we can see that EWR has a 8.78 min average delay, JFK has a 3.93 average delay, and JFK actually has a negative "average delay", so JFK on average actually arrives 2.38 minutes before arrival

```{r plot the data}

ggplot(data = delta_airline, aes(y= avg_delay, x= origin, fill = origin)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_brewer(palette = "Dark2") +
  labs(y="Average Arrival Delay", x= "Origin Airport", title= "Which origin airport has the best average delay?", subtitle = "(When flying Delta Airlines)") +
  theme_light()


```


QUESTION 3:
Which destination airport is the worst (you decide on the metric for worst) airport for arrival time?

```{r filter data}

worst_delay <- flights %>%
  select(arr_delay, dest) %>%
  drop_na() %>%
  group_by(dest) %>%
  summarise(avg_delay = mean(arr_delay)) %>%
  arrange(desc(avg_delay))

worst_delay


```
I chose to decide which destination airport was the worst based on the average arrival delay to that airport.  You can see that a destination of CAE (an airport in Lexington County, South Carolina) has an average arrival delay of nearly 42 minutes.  That is a significant (and the worst) delay of all destination airports. 

```{r make a visual}
top_worst_delays <- worst_delay %>%
  filter(avg_delay >= 18)

ggplot(data = top_worst_delays, aes(y= avg_delay, x= dest, fill = dest)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label=round(avg_delay)), nudge_y = 2) +
  scale_fill_brewer(palette = "Paired") +
  labs(y="Average Arrival Delay", x= "Destination Airport", title= "Which destination airport has the worst arrival time?", subtitle = "Top 10 airports with the worst arrival times") +
  theme_light()


```

